Distrivia: A Distributed Trivia Game

### 2011-04-08 papers/phase-one/paper.tex
Changes to paper

\section {What We've Done}
So far, our work has mainly focused on getting everything setup to work as a
distributed system. Getting the ground work done (for example servers and load
distributers running) has provided us a way to quickly develop the clients
and server as well as provided us with a way to reliably connect the servers to
the clients.
\subsection {Servers}
Currently, we have 6 servers running. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load distributers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly.
With this configuratoin if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.
\subsection {Traffic}
We have two load distributers set up. Both distributers are currently configured to
divide the requests among the 3 active servers. The duplicate distributers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load distributers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load distributers are set up the attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.

@@

\subsection {Messages}
We are currently using https protocol for our messages. This provides a secure
and usable messaging system that we have designed our client and server APIs to use.
The actual messages are sent and received in two way's. Initially a user session key is
genereated using the UUID \cite{uuid} algorithm and returned to the user on sucessful
authenticaton. This session key is then passed back to the server with every in coming
message. Messages are sent to the server using POST and GET request methods supported
by the HTTP protocol. Responses from the server are in the form of JSON \cite{json} objects.
JSON is a small, easily parsable format, representing objects in a human readable format as
they would be represented in JavaScript. With the web client the JSON object returned from
the server is instantly useable, however for the android client we have to parse the JSON object
into a native java object. The Android SDK provides libraries to do this, located in the
org.json \cite{orgjson} package namespace.
\section {What's Left}
\subsection {Web Client}
A lof of the main features of the web client are completed.
However, there are stil a few remaining features that need to be implemented.
The first is private games.
Joining them has only been stubbed out with no functionality.
The game playing routinees from public games should be compatible with private games, but may need a few tweeks as well.
Next, while it currently supports a universal leaderboard, it also needs to be able to display one ingame.
With the recent API changes, there are a few existing features which need to have thir API calls adjusted slightly.
Another feature that needs to be finished is displaying useful warnings and error messages.
While progress has been made on them and a reporting system has been built, messages still need to be fed into it.
Lastly, when a user clicks on a button to do something like log in, the button should be disabled while waiting for a response, alternatively it needs to be able to handle sending out two calls for the same request and not getting confused if it gets two back.
\subsection {Android}
So far, the only connection to the server that the android client has is when it
registers/logs in. We still need to connect the rest of the android client to
the server API to make sure all the communication between the android client
and the server are working. This includes joining public/private rounds,
competing in these rounds, and viewing leaderboards. While the UI is nearly
finished on the android client, nearly all of the connections to the servers
has yet to be done. Additionally, the android client needs to have more user
feedback and increased ability for the users to control which activity they
are in.
\subsection {Server}
Private game joining
Global leaderboard

\section {Meeting the Requirements}

Our system has two unique clients. One of them is designed for the Android platform as an app.
The other is a web client which is specifically being tested on Desktop and Laptop
browsers, however it may work on many handheld web browsers as well.

@@

We use round robin DNS to allow our clients to be completely ignorant of our servers.  Load distributors
can be added and removed from the DNS server and the clients will be completely ignorant of this
and unaffected.  The load distributors in turn know aboutour data servers and do their best to
distribute the load across them.  When a new server is added, it is just added to our load distributor
configuration file and clients can start using it.  When a server fails, the distributors will stop directing
traffic to it.

We are currently using Riak as our database backend. We chose Riak for multiple
reasons, but the biggest was it's focus on  scalability, fault-tolerance,
availability, and replication. Riak is a Amazon Dynamo \cite{dynamo} inspired
database, it supports full masterless replication. In Riak every node can
server client requests and they are propogated to the rest of the cluster.
In our system we have made sure that this replication occurs's between nodes
on Amazon EC2's internal Gigabit ethernet so it is as fast as possible. Riak
also has the unique property that if all of the nodes but one fail, the single
node can still accept both reads and writes to the database. When the other
nodes return to operational state from failure, all changes will automatically
be propogated to all of the nodes. This feature is the basis for our ability to
support and recorver from partial failures. In the event that two Riak nodes accept
writes from two seperate clients modifying the same data and Riak is unable
to resolve the conflict the user is given both versions and allowed to resolve
the conflict them selves. However we have configured our installation so this
can not occur, we have specified that every available node must agree on a
write before it is commited. Given the relatively small number of nodes (five)
that will take part in this agreement, we decided that the possible overhead
was acceptable for the increased level of consistency it gives us.

>>>>

\section {What We've Done}
So far, our work has mainly focused on getting everything setup to work as a
distributed system. Getting the ground work done (for example servers and load
sharers running) has provided us a way to quickly develop the clients
and server as well as provided us with a way to reliably connect the servers to
the clients.
\subsection {Servers}
Currently, we have 6 servers running. One server is being used as a web-monitor
host. We are able to login to the web-monitor to see the status of all our
servers, databases, and load sharers. This will help us manage if the services
are running and allow us to track down any problems, if they arise. Three
servers are all running our server application and hosting the databases.
We also have two final servers that we have setup as hot spares in the event
of partial or total failure.
Each server contains the server application that responds to clients as well as
the Riak \cite{riak} databases. Riak auto-synchronizes the databases between the 5 servers.
Note that the hot spares are setup to be part of the database cluster so they
have a full copy of the database on hand, however they don't accept any reads
or writes directly.
With this configuration if we were to add a server, we would just need to update 
Riak and it would sync the databases to any additional servers as well. These servers,
therefore, all contain the same information. If any servers go down, the distributed
system would still run fine. We could have all servers go down except one and still
be reliably handling requests.
\subsection {Traffic}
We have two load sharers set up. Both sharers are currently configured to
divide the requests among the three active servers. The duplicate sharers allow
for a load distributer to go down and we would still be able to serve requests
to the servers. These load sharers are selected by the clients using round
robin DNS where distrivia.lame.ws will select either load distributer that is
up at random. This provides us will stability for access to the servers. The
load sharers are setup to attempt to connect to a server twice before
deeming it dead and removing it from its list of servers to direct clients towards.

@@

\subsection {Messages}
We are currently using the https protocol for our messages. This provides a secure
and usable messaging system that we have designed our client and server APIs to use.
The actual messages are sent and received in two way's. Initially a user session key is
generated using the UUID \cite{uuid} algorithm and returned to the user on successful
authentication. This session key is then passed back to the server with every in coming
message. Messages are sent to the server using POST and GET request methods supported
by the HTTP protocol. Responses from the server are in the form of JSON \cite{json} objects.
JSON is a small, easily parseable format, representing objects in a human readable format as
they would be represented in JavaScript. With the web client the JSON object returned from
the server is instantly usable, however for the android client we have to parse the JSON object
into a native Java object. The Android SDK provides libraries to do this, located in the
org.json \cite{orgjson} package name space.
\section {What's Left}
\subsection {Server}
The server is currently almost feature complete, but still needs a lot of
testing. The only features it does not currently support is any sort of notion
of a private game. Client registration, and login are all supported. A client can
join a game, retrieve questions for that game, and answer questions. Both a
global leader board and individual per game leader board can also be requested.
As a group we have already found several inconsistencies and things we would
like to change with how server API requests are structured. We have also thought of
way's we can combine multiple requests into one, reducing the possible latency
and overall communication time of multiple requests. To fix these
inconsistencies members of the group worked to create a formal specification
for the request protocol of each possible request. This allows the server and client
programmers to work towards a common goal with out either making inappropriate
assumptions about how the system will work. We are currently in the middle
of refactoring the server to use the new requests detailed in the API.

\subsection {Web Client}
A lot of the main features of the web client are completed.
However, there are still a few remaining features that need to be implemented.
The first is private games.
Joining them has only been stubbed out with no functionality.
The game playing routines from public games should be compatible with private games, but may need a few tweaks as well.
Next, while it currently supports a universal leader board, it also needs to be able to display one in game.
With the recent API changes, there are a few existing features which need to have their API calls adjusted slightly.
Another feature that needs to be finished is displaying useful warnings and error messages.
While progress has been made on them and a reporting system has been built, messages still need to be fed into it.
Lastly, when a user clicks on a button to do something like log in, the button should be disabled while waiting for a response,
alternatively it needs to be able to handle sending out two calls for the same request and not getting confused if it gets two back.
\subsection {Android}
So far, the only connection to the server that the android client has is when it
registers/logs in. We still need to connect the rest of the android client to
the server API to make sure all the communication between the android client
and the server are working. This includes joining public/private rounds,
competing in these rounds, and viewing leader boards. While the UI is nearly
finished on the android client, nearly all of the connections to the servers
has yet to be done. Additionally, the android client needs to have more user
feedback and increased ability for the users to control which activity they
are in.
\section {Meeting the Requirements}

Our system has two unique clients. One of them is designed for the Android platform as an app.
The other is a web client which is specifically being tested on Desktop and Laptop
browsers, however it may work on many hand held web browsers as well.

@@

We use round robin DNS to allow our clients to be completely ignorant of our servers.  Load distributors
can be added and removed from the DNS server and the clients will be completely ignorant of this
and unaffected.  The load sharers in turn know about our data servers and do their best to
distribute the load across them.  When a new server is added, it is just added to our load distributor
configuration file and clients can start using it.  When a server fails, the distributors will stop directing
traffic to it.

We are currently using Riak as our database backend. We chose Riak for multiple
reasons, but the biggest was it's focus on  scalability, fault-tolerance,
availability, and replication. Riak is a Amazon Dynamo \cite{dynamo} inspired
database, it supports full master less replication. In Riak every node can
server client requests and they are propagated to the rest of the cluster.
In our system we have made sure that this replication occurs between nodes
on Amazon EC2's internal Gigabit ethernet so it is as fast as possible. Riak
also has the unique property that if all of the nodes but one fail, the single
node can still accept both reads and writes to the database. When the other
nodes return to operational state from failure, all changes will automatically
be propagated to all of the nodes. This feature is the basis for our ability to
support and recover from partial failures. In the event that two Riak nodes accept
writes from two separate clients modifying the same data and Riak is unable
to resolve the conflict the user is given both versions and allowed to resolve
the conflict them selves. However we have configured our installation so this
can not occur, we have specified that every available node must agree on a
write before it is committed. Given the relatively small number of nodes (five)
that will take part in this agreement, we decided that the possible overhead
was acceptable for the increased level of consistency it gives us.
